---
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/WORK/kings")
```

```{r, include=FALSE}
library(dotenv)
library(tidyverse)
library(stargazer)
library(ggraph)
library(igraph)
library(migraph)
library(data.table)
library(ggcorrplot)

setwd("~/Documents/WORK/kings")
load_dot_env()

place_existance <- readRDS("EJ_DAC_Paper/Data/place_existance.RDS")
all_places <- bind_rows(place_existance)
network_fp <- paste0(Sys.getenv("BOX_PATH"), "/EJ_Paper/cleaned_extracts_DACified")
extract_list = list.files(network_fp)
gsp_ids <- gsub("^0+", "", gsub("\\.RDS", "", extract_list))

all_nodes <- data.frame()

for (g in seq_along(gsp_ids)) {
   net <- readRDS(paste0(network_fp, "/", extract_list[g]))
   gsp_id <- paste0("gsp_",gsp_ids[g])
   nodes <- tibble(igraph::as_data_frame(net, what = "vertices"))
   places <- nodes %>% filter(exists == 1)
   nodes$gsp_id <- gsp_id
   mentions <- rowSums(nodes %>% 
                          select(admin_sum, 
                                 basin_plan_sum, 
                                 sust_criteria_sum,
                                 monitoring_networks_sum,
                                 projects_mgmt_actions_sum))
   nodes <- nodes %>% 
      rename(admin= admin_sum, 
             basin_plan = basin_plan_sum,
             sust_criteria = sust_criteria_sum,
             monitoring_networks = monitoring_networks_sum, 
             projects_mgmt_actions = projects_mgmt_actions_sum) %>% 
      mutate(admin_per = 100*(admin/mentions),
             basin_plan_per = 100*(basin_plan/mentions),
             sust_criteria_per = 100*(sust_criteria/mentions),
             monitoring_networks_per = 100*(monitoring_networks/mentions),
             projects_mgmt_actions_per = 100*(projects_mgmt_actions/mentions),
             eig_std = (eigenvector - mean(eigenvector, na.rm=TRUE)) / sd(eigenvector, na.rm=TRUE),
             leader_std = (leader_dist_min - mean(leader_dist_min, na.rm=TRUE)) / sd(leader_dist_min, na.rm=TRUE),
             deg_std = (degree - mean(degree, na.rm=TRUE)) / sd(degree, na.rm=TRUE),
             leader_dist_min = ifelse(is.infinite(leader_dist_min), NA, leader_dist_min),
             MHI_std = MHI/10000,
             POP_std = POP/1000000,
             is_place = ifelse(is.na(GEOID20), 0, 1),
             )
   all_nodes <- rbind(all_nodes, nodes)
}

all_place_nodes <- all_nodes %>% 
   filter(is_place == 1) %>% 
   mutate(eig_std_2 = (eigenvector - mean(eigenvector, na.rm=TRUE)) / sd(eigenvector, na.rm=TRUE),
          appears_in_sec = ifelse(is.nan(admin_per), 0, 1))
   
```

```{r}

network_fp <- paste0(Sys.getenv("BOX_PATH"), 
                     "/EJ_Paper/cleaned_extracts_DACified_substantive")
extract_list = list.files(network_fp)
gsp_ids <- gsub("^0+", "", gsub("\\.RDS", "", extract_list))

sub_all_nodes <- data.frame()

for (g in seq_along(gsp_ids)) {
   net <- readRDS(paste0(network_fp, "/", extract_list[g]))
   gsp_id <- paste0("gsp_",gsp_ids[g])
   nodes <- tibble(igraph::as_data_frame(net, what = "vertices"))
   places <- nodes %>% filter(exists == 1)
   nodes$gsp_id <- gsp_id
   mentions <- rowSums(nodes %>% 
                          select(admin_sum, 
                                 basin_plan_sum, 
                                 sust_criteria_sum,
                                 monitoring_networks_sum,
                                 projects_mgmt_actions_sum))
   nodes <- nodes %>% 
      rename(admin= admin_sum, 
             basin_plan = basin_plan_sum,
             sust_criteria = sust_criteria_sum,
             monitoring_networks = monitoring_networks_sum, 
             projects_mgmt_actions = projects_mgmt_actions_sum) %>% 
      mutate(admin_per = 100*(admin/mentions),
             basin_plan_per = 100*(basin_plan/mentions),
             sust_criteria_per = 100*(sust_criteria/mentions),
             monitoring_networks_per = 100*(monitoring_networks/mentions),
             projects_mgmt_actions_per = 100*(projects_mgmt_actions/mentions),
             eig_std = (eigenvector - mean(eigenvector, na.rm=TRUE)) / sd(eigenvector, na.rm=TRUE),
             leader_std = (leader_dist_min - mean(leader_dist_min, na.rm=TRUE)) / sd(leader_dist_min, na.rm=TRUE),
             deg_std = (degree - mean(degree, na.rm=TRUE)) / sd(degree, na.rm=TRUE),
             leader_dist_min = ifelse(is.infinite(leader_dist_min), NA, leader_dist_min),
             MHI_std = MHI/10000,
             POP_std = POP/1000000,
             is_place = ifelse(is.na(GEOID20), 0, 1),
             )
   sub_all_nodes <- rbind(sub_all_nodes, nodes)
}

sub_all_place_nodes <- sub_all_nodes %>% 
   filter(is_place == 1) %>% 
   mutate(eig_std_2 = (eigenvector - mean(eigenvector, na.rm=TRUE)) / sd(eigenvector, na.rm=TRUE))
   
```


###
### HYPOTHESIS 1a: DACs have lower existence than non-disadvantaged communities
###

```{r}
#cross-tab for (DAC, exists)
with(all_places, table(exists, DAC))

# glm for if DAC influences exists
exists_mod <- glm(exists~DAC+POP+incorporated+per_latino, 
                family = binomial,
                data = all_places)

exists_mod_2 <- glm(exists ~ MHI+POP+incorporated+per_latino, 
                  family = binomial,
                  data = all_places)

stargazer(exists_mod, exists_mod_2, type='text')

stargazer(exists_mod, exists_mod_2, type='html', out = 'EJ_DAC_Paper/Out/mods/0_exists_mod.html')
```


###
### HYPOTHESIS 1b: DACs have lower count within sections than non-disadvantaged communities
###

```{r, results='asis'}

a_dac <- glm(admin ~ DAC+
               POP_std+
               incorporated+
               per_latino,
             family=poisson,
              data = all_place_nodes)


b_dac <- glm(basin_plan ~ DAC+
               POP_std+
               incorporated+
               per_latino,
             family=poisson,
              data = all_place_nodes)

c_dac <- glm(sust_criteria ~ DAC+
               POP_std+
               incorporated+
               per_latino,
             family=poisson,
              data = all_place_nodes)

d_dac <- glm(monitoring_networks ~ DAC+
               POP_std+
               incorporated+
               per_latino,
             family=poisson,
              data = all_place_nodes)

e_dac <- glm(projects_mgmt_actions ~ DAC+
               POP_std+
               incorporated+
               per_latino,
             family=poisson,
              data = all_place_nodes)

stargazer(a_dac, b_dac, c_dac, d_dac, e_dac, type='text')

stargazer(a_dac, b_dac, c_dac, d_dac, e_dac, type='html', out = 'EJ_DAC_Paper/Out/mods/6_dac_locs_mod.html')

```

###
### HYPOTHESIS 2a: DACs have lower eigenvector centrality than non-disadvantaged communities
###
### HYPOTHESIS 2b: DACs have lower eigenvector centrality than non-disadvantaged communities in substantive sections
###

```{r, results='asis'}

eig_mod <- lm(log(eig_std) ~ MHI_std+
                 POP_std+
                 incorporated+
                 per_latino,
              data = all_place_nodes)

eig_mod_2 <- lm(log(eig_std) ~ DAC+
                 POP_std+
                 incorporated+
                 per_latino,
              data = all_place_nodes)

eig_mod_3<- lm(log(eig_std) ~ MHI_std+
                 POP_std+
                 incorporated+
                 per_latino,
              data = sub_all_place_nodes)

eig_mod_4 <- lm(log(eig_std) ~ DAC+
                 POP_std+
                 incorporated+
                 per_latino,
              data = sub_all_place_nodes)

stargazer(eig_mod, eig_mod_2, eig_mod_3, eig_mod_4, type='text')

stargazer(eig_mod, eig_mod_2, eig_mod_3, eig_mod_4, type='html', 
          out = 'EJ_DAC_Paper/Out/mods/1_eig_mod.html')
```


###
### HYPOTHESIS 3a: DACs have lower leader closeness than non-disadvantaged communities
###
### HYPOTHESIS 3b: DACs have lower leader closeness than non-disadvantaged communities in substantive sections
###

```{r, results='asis'}

lead_mod <- glm(leader_dist_min ~ MHI_std+
                 POP_std+
                 incorporated+
                 per_latino,
                family=poisson,
              data = all_place_nodes)


lead_mod_2 <- glm(leader_dist_min ~ DAC+
                 per_latino+
                 POP_std+
                 incorporated,
                 family=poisson,
              data = all_place_nodes)

lead_mod_3 <- glm(leader_dist_min ~ MHI_std+
                 POP_std+
                 incorporated+
                 per_latino,
                family=poisson,
              data = sub_all_place_nodes)


lead_mod_4 <- glm(leader_dist_min ~ DAC+
                 per_latino+
                 POP_std+
                 incorporated,
                 family=poisson,
              data = sub_all_place_nodes)

stargazer(lead_mod, lead_mod_2, lead_mod_3, lead_mod_4, type='text')
          
stargazer(lead_mod, lead_mod_2, lead_mod_3, lead_mod_4, type='text',
          out = 'EJ_DAC_Paper/Out/mods/2_lead_mod.html')
```

###
### ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
### NOT CURRENTLY IN USE
###

```{r, results='asis'}

eig_a <- lm(log(eig_std) ~ admin_per,
              data = all_nodes)


eig_b <- lm(log(eig_std) ~ basin_plan_per,
              data = all_nodes)


eig_c <- lm(log(eig_std) ~ sust_criteria_per,
              data = all_nodes)


eig_d <- lm(log(eig_std) ~ monitoring_networks_per,
              data = all_nodes)


eig_e <- lm(log(eig_std) ~ projects_mgmt_actions_per,
              data = all_nodes)

stargazer(eig_a, eig_b, eig_c, eig_d, eig_e, type='text')

stargazer(eig_a, eig_b, eig_c, eig_d, eig_e, type='html', out = 'EJ_DAC_Paper/Out/mods/3_eig_locs_mod.html')
```


```{r, results='asis'}

lead_a <- glm(leader_dist_min ~ admin_per,
              family=poisson,
              data = all_nodes)


lead_b <- glm(leader_dist_min ~ basin_plan_per,
              family=poisson,
              data = all_nodes)


lead_c <- glm(leader_dist_min ~ sust_criteria_per,
              family=poisson,
              data = all_nodes)


lead_d <- glm(leader_dist_min ~ monitoring_networks_per,
              family=poisson,
              data = all_nodes)


lead_e <- glm(leader_dist_min ~ projects_mgmt_actions_per,
              family=poisson,
              data = all_nodes)

stargazer(lead_a, lead_b, lead_c, lead_d, lead_e, type='text')

stargazer(lead_a, lead_b, lead_c, lead_d, lead_e, type='html', out = 'EJ_DAC_Paper/Out/mods/4_lead_locs_mod.html')

```

```{r, results='asis'}

a_dac <- glm(admin ~ is_place,
             family = poisson,
              data = all_nodes)


b_dac <- glm(basin_plan ~ is_place,
             family = poisson,
              data = all_nodes)

c_dac <- glm(sust_criteria ~ is_place,
             family = poisson,
              data = all_nodes)

d_dac <- glm(monitoring_networks ~ is_place,
             family = poisson,
              data = all_nodes)

e_dac <- glm(projects_mgmt_actions ~ is_place,
             family = poisson,
              data = all_nodes)

stargazer(a_dac, b_dac, c_dac, d_dac, e_dac, type='text')

stargazer(a_dac, b_dac, c_dac, d_dac, e_dac, type='html', out = 'EJ_DAC_Paper/Out/mods/5_place_locs_mod.html')
```
